A Survey of Quantitative Descriptions of Molecular Structure

### 2012-04-09 paper.tex
Started adding impl table

@@

Given the multiple implementations of many descriptors, it is natural
to compare their performance. As noted previously, it is desirable,
but unlikely that implementations of the same descriptor using
different cheminformatcs toolkits will be identical. Gupta et al
employed SMARTS based descriptors from the CDK and MOE to develop
decision tree models to predict human liver microsomal metabolic
stability\cite{Gupta:2010uq}. Their results indicated very similar
performance between the two implementations.

@@

A number of physicochemical descriptors commonly employed in
predictive modeling are themselves models of an experimental
property. For example, log P can be experimentally measured and a
number of algorithms have been developed to predict log P
[REFs]. Given the utility of log P in drug discovery, a number of
implementations are available. Given that this descriptor is a
surrogate for an actual experimental property, it is reasonable to
compare calculated log P values from different implementations to the
experimentall observed values, rather than between themselves. Figure
\ref{fig:logp} compares computed log P values from the CDK, ChemAxon
and ACD Labs for a set of 10,000 molecules taken from the logPstar
dataset [REF].

@@

\section{Running Notes}
\begin{itemize}
\item \textbf{Done} What are descriptors, why are they needed?
\item Survey of descriptor types, focusing on nature of inputs,
  complexity, interpretabilty 
\item \textbf{Done} Talk about biological descriptors (see paper from Tropsha that
  uses bioactivity data as descriptors)
\item \textbf{Done} Discuss how fingerprints can also be used as descriptors
\item Also mention descriptors for non-small molecule cases
  \begin{itemize}
  \item proteins
  \item polymers \& mixtures
  \item \textbf{Done} inorganics (maybe ask Maciej for pointers)
  \end{itemize}
\item Survey of descriptor implementations (OSS and non-OSS)
  \begin{itemize}
  \item MOE
  \item DRAGON
  \item CDK
  \item RDKit
  \item JChem
  \item PipelinePilot
  \item OEChem/GraphTK
  \item JOELib
  \item CODESSA
  \item Schrodinger stuff (Meastro?)
  \end{itemize}
\item Go into more detail of OSS descriptor implementations
  \begin{itemize}
  \item CDK (describe architecture), mention wrappers around the
    library (CDKDescUI and batch mode)
  \item PaDEL
  \end{itemize}
\item Compare OSS descs to commercial ones - point to the Pfizer/CDD
  paper that compared MOE and CDK
\end{itemize}
\clearpage
\newpage

@@

>>>>

Given the focus of this article on Open Source implementations, we
have noted the availability in Table \ref{tab:impl}. While a number of
implementations are not strictly Open Source according to OSI
definitions, the fact that they provide free academic licenses does
allow them to be used somewhat freely.

@@

Given the multiple implementations of many descriptors, it is natural
to compare their performance. As noted previously, it is desirable,
but unlikely, that implementations of the same descriptor using
different cheminformatcs toolkits will be identical. Gupta et al
employed SMARTS based descriptors from the CDK and MOE to develop
decision tree models to predict human liver microsomal metabolic
stability\cite{Gupta:2010uq}. Their results indicated very similar
performance between the two implementations.

@@

A number of physicochemical descriptors commonly employed in
predictive modeling are themselves models of an experimental
property. For example, log P can be experimentally measured and a
number of algorithms have been developed to predict $\log P$
\cite{Wang:1997aa,Wildman:1999aa}. Given the utility of log P in drug
discovery, a number of implementations are available. Given that this
descriptor is a surrogate for an actual experimental property, it is
reasonable to compare calculated log P values from different
implementations to the experimentally observed values, rather than
between themselves. Figure \ref{fig:logp} compares computed log P
values from the CDK, ChemAxon and ACD Labs for a set of 10,000
molecules taken from the logPstar dataset [REF].

@@

\section{Running Notes}
\begin{itemize}
\item Survey of descriptor implementations (OSS and non-OSS)
  \begin{itemize}
  \item MOE
  \item DRAGON
  \item CDK
  \item RDKit
  \item JChem
  \item PipelinePilot
  \item OEChem/GraphTK
  \item JOELib
  \item CODESSA
  \item Schrodinger stuff (Meastro?)
  \end{itemize}
\item Go into more detail of OSS descriptor implementations
  \begin{itemize}
  \item CDK (describe architecture), mention wrappers around the
    library (CDKDescUI and batch mode)
  \item PaDEL
  \end{itemize}
\item Compare OSS descs to commercial ones - point to the Pfizer/CDD
  paper that compared MOE and CDK
\end{itemize}
\clearpage
\newpage

@@

%% See http://code.google.com/p/rdkit/w/list for RDKit
\ctable[caption={A summary of descriptor implementations covering Open
  Source and commercial cases.}, cap={}, label={tab:impl}]
{llllp{0.3\linewidth}p{0.2\linewidth}}
{}
{
\FL
Name & License & Toolkit & Application & Descriptors & Fingerprints
\ML
CDK & LGPL & Yes & \raggedright{CDKDescUI} &
\raggedright{Constitutional, topological, geometric, electronic and hybrid} &
  \raggedright{MACCS, PubChem, Kier \& Hall Estate keys. Hashed linear
  and circular} \NN
JChem & Commercial & Yes & cxcalc & \raggedright{} & raggedright{} \NN
JOELib & GPL & Yes & & \raggedright{} & \raggedright{} \NN
OpenBabel & GPL & Yes & & \raggedright{} & \raggedright{} \NN
RDKit & New BSD & Yes & & \raggedright{Constitutional, topological,
  EState, partial charges and VSA combining EState, partial charges
  and MR} & MACCS, hashed, atom pairs, topological
torsions 
\LL
}
